---
title: "Using R with MySQL"
author: "Jasper Ginn"
date: "19 May 2015"
output: html_document
---

# Introduction

This document briefly outlines how you can use R to access MySQL databases.

```{r}
# Set global knitr Options
require(knitr)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, tidy = FALSE, results = 'asis')
```

# Connecting to the database

Firstly, make sure that MySQL is running by opening up a terminal and executing the following line:

 - sudo /usr/local/mysql/support-files/mysql.server start

In R, you need to install the "RMySQL" package, which you can do as follows

```{r}
# Install packages
install.packages("RMySQL")
# Load
library(RMySQL)
```

Assuming you have already stored data in a MySQL database, you can connect to a table like this:

```{r}
# Specify credentials
usern = "root"
dbname = "terror001"
password = ""
# Connect
con = dbConnect(MySQL(), username=usern, dbname=dbname, password = password)
```

It is straightforward to get the tables residing in a database

```{r}
# List tables
dbListTables(con)
```

You can also list the fields in a table:

```{r}
# List fields
dbListFields(con, 'forum_comments')
```

There are several ways to select a table. For small tables, you can query the entire table and process in R, like so:

```{r}
# Read table
tab <- dbReadTable(con, "activity_log")
```

However, for larger tables you might want to use SQL statements before quering the data:

```{r}
# Read part of a table
tab <- dbGetQuery(con, "SELECT * FROM forum_comments")
```

Don't forget to disconnect!

```{r}
# Disconnect
dbDisconnect(con)
```

# More resources:

 1. [Accessing a Database from R](http://www.stat.berkeley.edu/~nolan/stat133/Fall05/lectures/SQL-R.pdf)
 2. [RMySQL documentation](http://cran.r-project.org/web/packages/RMySQL/RMySQL.pdf)
 3. [RMySQL basics](https://mkmanu.wordpress.com/2014/07/24/r-and-mysql-a-tutorial-for-beginners/)
 4. [In case installing RMySQL on Windows gives you issues](http://stackoverflow.com/questions/5223113/using-mysql-in-r-for-windows)